### This code produces time series plots of steps taken by activity monitoring device users  
### Author: Aaron Levine  

#### Load libraries  
```{r}
library(dplyr)
library(plyr)
```


#### read in csv file  
```{r}
activity <- read.csv("activity.csv")
```

#### remove missing values  
```{r}
activitynonan <- filter(activity,!is.na(activity$steps))
```

#### mean number of steps, grouped by date  
```{r}
stepsperday<-tapply(activitynonan$steps,activitynonan$date,mean)
```

#### remove dates with no steps  
```{r}
stepsperday<-stepsperday[!is.na(stepsperday)]
```

#### make histogram  
```{r echo=TRUE}
hist(stepsperday,xlab="Steps Per Day",ylab="Number of Days", main="Histogram of Mean Steps Per Day")
```

#### mean steps per day  
```{r echo=TRUE}
mean(stepsperday)
```

#### median steps per day  
```{r echo=TRUE}
median(stepsperday)
```

#### create time series of steps per interval, each day, on average  
```{r warning=FALSE}
timeseriessteps<-aggregate(activitynonan,list(activitynonan$interval),mean)

plot(timeseriessteps$interval,timeseriessteps$steps,type="l",xlab="Interval (5 minutes)", ylab="Mean Number of Steps")
```

#### time of max steps per day  
```{r echo=TRUE}
timeseriessteps[timeseriessteps$steps==max(timeseriessteps$steps),]$interval
```

#### function to remove missing values and replace with interval averages  
```{r echo=TRUE}
removena <- function(steps, interval,meansteps) {
  if (is.na(steps)) {
    steps<-meansteps[meansteps$interval==interval,]$steps}
  steps
}
```

#### remove missing values by applying removena function  
```{r echo=TRUE}
activity['stepsnonan'] <- apply(activity[,c('steps','interval')],1,function(x) removena(x[1],x[2],timeseriessteps))
```

#### function to add isweekend feature  
```{r echo=TRUE}
setweekend <- function(date){
  dayofweek <- weekdays(as.POSIXlt(as.character(date),format="%Y-%m-%d"))
  isweekend <- "weekday"
  if(dayofweek=='Saturday'|dayofweek=='Sunday'){
    isweekend<-"weekend"
  }
  isweekend
}
```

#### apply function to add isweekend feature  
```{r echo=TRUE}
activity['isweekend'] <- as.factor(sapply(activity[,c('date')],function(x) setweekend(x)))

```
#### make time series for weekend and weekday and plot
```{r warning=FALSE}
timeseriessteps_weekend<-aggregate(activity[activity$isweekend=="weekend",],list(activity[activity$isweekend=="weekend",]$interval),mean)
timeseriessteps_weekday<-aggregate(activity[activity$isweekend=="weekday",],list(activity[activity$isweekend=="weekday",]$interval),mean)

par(mfrow=c(2,1))
plot(timeseriessteps_weekday$interval,timeseriessteps_weekday$stepsnonan,type="l",xlab="Interval (5 minutes)", ylab="Mean Number of Steps",main="Weekday")
plot(timeseriessteps_weekend$interval,timeseriessteps_weekend$stepsnonan,type="l",xlab="Interval (5 minutes)", ylab="Mean Number of Steps",main="Weekend")
```
